Case Study Reasoning about BYOD Policies in the Healthcare Sector

### 2016-06-30 paper-2.tex
justifies why apppal

The \ac{HiMMS} policy is taken from their \emph{Mobile Security Toolkit} for assisting healthcare organisations and security practitioners in managing the security of their mobile computing devices. 
\ac{HiMMS} itself is an American non-profit company advising healthcare companies on IT.
Their policy~\cite{healthcare_information_and_management_systems_society_mobile_2012} takes the form of a short user agreement that it's subjects are expected to sign.  
This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
It also congtains many rules where the subject is expected to acknowledge a risk or regulation.  
For example in one rule the subject is required to acknowledge that networks can be unreliable:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMMS}}:~\itshape 
I understand and accept that synchronisation relies on one or more cellular network providers and the Internet, and that both are subject to slowdowns and outages of extended duration that are beyond the control of IT.}]
'himms' says User can-say User hasAcknowledged('internet-unreliable')
  if User isUser.

@@

The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series fo \emph{trusts}, each of which contains several hospitals.  A doctor may work at several hospitals within a trust.}. 
It is interesting in part because it represents a \ac{BYOD} policy in actual use, but also because it is significantly more complex than most \ac{BYOD} policies as it goes beyond basic requirements such as requiring encryption by describing relationships between organisational groups within the trust and precise organisation specific requirements.
Like the \ac{HiMMS} policy it has rules that just require an acknowledgement (for example Rule~3.2 requires that the device owner \emph{``accepts that the device can be used to communicate through all corporate channels''}) but there is a greater emphasis on the specifying the responsibilities of different groups and roles within the hospital.  For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{NHS}}:~\itshape 
Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
'nhs-trust' says Staff hasAppliedForPhone(Form)
  if Staff hasSubmitted(Form),
     'assistant-director' hasAuthorized(Form),
     'finance-director' hasAuthorized(Form).

@@

The SANS Institute is a private US company that specialise in information security.
Whilst their example \ac{BYOD} policy is not healthcare specific, it is still interesting as there advice is widely regarded, and contains more general security advice.
The SANS \ac{BYOD} policy is split into sections that describe how device features, such as Bluetooth or OTA provisioning, should be configured.  Some rules are acknowledgements: for example the use of internet services in Section~2.8 requires acknowledgement of the email security, acceptable use and an eCommerce security policy.
Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
For example in Section~2.4 the SANS policy states that infrared support should be disabled if Blutooth is available: this is easilly expressed in AppPAL:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
'sans' says Device shouldDisable('irda')
  if Device hasFeature('bluetooth').
\end{lstlisting}

@@

>>>>

The \ac{HiMMS} policy is taken from their \emph{Mobile Security Toolkit} for assisting healthcare organisations and security practitioners in managing the security of their mobile computing devices. 
\ac{HiMMS} itself is an American non-profit company advising healthcare companies on IT.
Their policy~\cite{healthcare_information_and_management_systems_society_mobile_2012} takes the form of a short user agreement that it's subjects are expected to sign.  
This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
It also contains many rules where the subject is expected to acknowledge a risk or regulation.  
For example in one rule the subject is required to acknowledge that networks can be unreliable:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMMS}}:~\itshape 
I understand and accept that synchronisation relies on one or more cellular network providers and the Internet, and that both are subject to slowdowns and outages of extended duration that are beyond the control of IT.}]
'himms' says User can-say User hasAcknowledged('internet-unreliable')
  if User isUser.

@@

The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series of \emph{trusts}, each of which contains several hospitals.  A doctor may work at several hospitals within a trust.}. 
It is interesting in part because it represents a \ac{BYOD} policy in actual use, but also because it is significantly more complex than most \ac{BYOD} policies as it goes beyond basic requirements such as requiring encryption by describing relationships between organisational groups within the trust and precise organisation specific requirements.
Like the \ac{HiMMS} policy it has rules that just require an acknowledgement (for example Rule~3.2 requires that the device owner \emph{``accepts that the device can be used to communicate through all corporate channels''}) but there is a greater emphasis on the specifying the responsibilities of different groups and roles within the hospital.  For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{NHS}}:~\itshape 
Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
'nhs-trust' says Staff hasAppliedForPhone(Form)
  if Staff hasSubmitted(Form),
     'assistant-director' hasAuthorized(Form),
     'finance-director' hasAuthorized(Form).

@@

The SANS Institute is a private US company that specialise in information security.
Whilst their example \ac{BYOD} policy is not healthcare specific, it is still interesting as there advice is widely regarded, and contains more general security advice.
The SANS \ac{BYOD} policy is split into sections that describe how device features, such as Bluetooth or OTA provisioning, should be configured.  Some rules are acknowledgements: for example the use of internet services in Section~2.8 requires acknowledgement of the email security, acceptable use and an eCommerce security policy.
Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
For example in Section~2.4 the SANS policy states that infrared support should be disabled if Bluetooth is available: this is easily expressed in AppPAL:
\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
'sans' says Device shouldDisable('irda')
  if Device hasFeature('bluetooth').
\end{lstlisting}

@@

\subsection{Why AppPAL?}

AppPAL is an authorization language designed to model app preferences~\cite{hallett_apppal_2016}.
It was created by instantiating the SecPAL language (which was designed for access control in distributed computing environments~\cite{becker_secpal:_2010}) with predicates and checking functions for describing apps.
It supports delegation, roles and constraints including arbitrary external checking programs to describe precisely what apps users might want to use.

AppPAL (and SecPAL) is designed to be readable.
It is powerful enough to describe complex policies and rules, whilst also being semantically simple having only three rules for evaluation.
When reasoning about \ac{BYOD} policies, it simplicity and readability gives it an advantage over more powerful languages such as XACML~\cite{oasis_extensible_2013} as a rule written in AppPAL can be comprehended quickly.
Whilst an end-user might not write a policy in a formal language, an IT administrator might wish to use a formal language to aid with reasoning and checking the policies they write.
AppPAL gives these IT professionals an easy entry point to formal methods, without over complicating things.
Early experiments showed that AppPAL can describe \ac{BYOD} policies, and help make clarify mistakes ahead of time~\cite{hallett_specifying_2016}.

An AppPAL evaluation library has been developed that runs on Java and Android and is available online\footnote{\url{https://github.com/apppal/libapppal}}.
As part of this work we have developed (in Section~\ref{sec:details}) a series of policy analysis tools for AppPAL.
These are compiled alongside the main AppPAL interpreter and may be run from the command-line.
